---
title: "In-Depth 1: Comparison of Point-Estimates"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, bayesian, posterior, test]
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{In-Depth 1: Comparison of Point-Estimates}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
csl: apa.csl
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)

if (!requireNamespace("see", quietly = TRUE) ||
  !requireNamespace("parameters", quietly = TRUE) ||
  !requireNamespace("ggplot2", quietly = TRUE) ||
  !requireNamespace("datawizard", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
}

options(knitr.kable.NA = "")
knitr::opts_chunk$set(
  echo = TRUE,
  comment = ">",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  dpi = 150
)
options(digits = 2)
```


This vignette can be referred to by citing the package:

- Makowski, D., Ben-Shachar, M. S., \& Lüdecke, D. (2019). *bayestestR: Describing Effects and their Uncertainty, Existence and Significance within the Bayesian Framework*. Journal of Open Source Software, 4(40), 1541. https://doi.org/10.21105/joss.01541

---

# Effect Point-Estimates in the Bayesian Framework


## Introduction 

One of the main difference between the Bayesian and the frequentist frameworks
is that the former returns a probability *distribution* for each effect (*i.e.*,
a model parameter of interest, such as a regression slope) instead of a *single
value*. However, there is still a need and demand - for reporting or use in
further analysis - for a single value (**point-estimate**) that best
characterises the underlying posterior distribution.

There are three main indices used in the literature for effect estimation: 
- the **mean** 
- the **median**
- the **MAP** (Maximum A Posteriori) estimate (roughly
corresponding to the mode - the "peak" - of the distribution)

Unfortunately, there is no consensus about which one to use, as no systematic
comparison has ever been done.

In the present work, we will compare these three point-estimates of effect with
each other, as well as with the widely known **beta**, extracted from a
comparable frequentist model. These comparisons can help us draw bridges and
relationships between these two influential statistical frameworks.


## Experiment 1: Relationship with Error (Noise) and Sample Size

### Methods

We will be carrying out simulation aimed at modulating the following
characteristics:

- **Model type**: linear or logistic.
- **"True" effect** (*known* parameters values from which data is drawn):
Can be 1 or 0 (no effect).
- **Sample size**: From 20 to 100 by steps of 10.
- **Error**: Gaussian noise applied to the predictor with SD uniformly spread
between 0.33 and 6.66 (with 1000 different values).

We generated a dataset for each combination of these characteristics, resulting
in a total of `2 * 2 * 9 * 1000 = 36000` Bayesian and frequentist models. The
code used for generation is available
[here](https://easystats.github.io/circus/articles/bayesian_indices.html)
(please note that it takes usually several days/weeks to complete).

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(datawizard)
library(see)
library(parameters)

df <- read.csv("https://raw.github.com/easystats/circus/master/data/bayesSim_study1.csv")
```


### Results


#### Sensitivity to Noise

```{r, message=FALSE, warning=FALSE}
dat <- df
dat <- data_select(dat, select = c("error", "true_effect", "outcome_type", "Coefficient", "Median", "Mean", "MAP"))
dat <- reshape_longer(
  dat,
  select = -c("error", "true_effect", "outcome_type"),
  names_to = "estimate",
  values_to = "value"
)
dat$temp <- as.factor(cut(dat$error, 10, labels = FALSE))

tmp <- lapply(split(dat, dat$temp), function(x) {
  x$error_group <- rep(round(mean(x$error), 1), times = nrow(x))
  return(x)
})

dat <- do.call(rbind, tmp)
dat <- data_filter(dat, value < 6)

ggplot(dat, aes(x = error_group, y = value, fill = estimate, group = interaction(estimate, error_group))) +
  # geom_hline(yintercept = 0) +
  # geom_point(alpha=0.05, size=2, stroke = 0, shape=16) +
  # geom_smooth(method="loess") +
  geom_boxplot(outlier.shape = NA) +
  theme_modern() +
  scale_fill_manual(
    values = c("Coefficient" = "#607D8B", "MAP" = "#795548", "Mean" = "#FF9800", "Median" = "#FFEB3B"),
    name = "Index"
  ) +
  ylab("Point-estimate") +
  xlab("Noise") +
  facet_wrap(~ outcome_type * true_effect, scales = "free")
```


#### Sensitivity to Sample Size

```{r, message=FALSE, warning=FALSE}

dat <- df
dat <- data_select(dat, select = c("sample_size", "true_effect", "outcome_type", "Coefficient", "Median", "Mean", "MAP"))
dat <- reshape_longer(
  dat,
  select = -c("sample_size", "true_effect", "outcome_type"),
  names_to = "estimate",
  values_to = "value"
)
dat$temp <- as.factor(cut(dat$sample_size, 10, labels = FALSE))

tmp <- lapply(split(dat, dat$temp), function(x) {
  x$size_group <- rep(round(mean(x$sample_size), 1), times = nrow(x))
  return(x)
})

dat <- do.call(rbind, tmp)
dat <- data_filter(dat, value < 6)

ggplot(dat, aes(x = size_group, y = value, fill = estimate, group = interaction(estimate, size_group))) +
  # geom_hline(yintercept = 0) +
  # geom_point(alpha=0.05, size=2, stroke = 0, shape=16) +
  # geom_smooth(method="loess") +
  geom_boxplot(outlier.shape = NA) +
  theme_modern() +
  scale_fill_manual(
    values = c("Coefficient" = "#607D8B", "MAP" = "#795548", "Mean" = "#FF9800", "Median" = "#FFEB3B"),
    name = "Index"
  ) +
  ylab("Point-estimate") +
  xlab("Sample size") +
  facet_wrap(~ outcome_type * true_effect, scales = "free")
```


#### Statistical Modelling

We fitted a (frequentist) multiple linear regression to statistically test the
the predict the presence or absence of effect with the estimates as well as
their interaction with noise and sample size.

```{r, message=FALSE, warning=FALSE}

dat <- df
dat <- data_select(dat, select = c("sample_size", "true_effect", "outcome_type", "Coefficient", "Median", "Mean", "MAP"))
dat <- reshape_longer(
  dat,
  select = -c("sample_size", "error", "true_effect", "outcome_type"),
  names_to = "estimate",
  values_to = "value"
)

out <- glm(true_effect ~ outcome_type / estimate / value, data = dat, family = "binomial")
out <- parameters(out, df_method = "wald")
out <- data_select(out, c("Parameter", "Coefficient", "p"))
out <- data_filter(out, grep("^outcome_type(.*):value$", x = out$Parameter))
out <- out[order(out$Coefficient, decreasing = TRUE), ]
knitr::kable(out, digits = 2)
```

This suggests that, in order to delineate between the presence and the absence
of an effect, compared to the frequentist's beta coefficient:

- For linear models, the **Mean** was the better predictor, closely followed by
the **Median**, the **MAP** and the frequentist **Coefficient**.
- For logistic models, the **MAP** was the better predictor, followed by the
**Median**, the **Mean** and, behind, the frequentist **Coefficient**.

Overall, the **median** appears to be a safe choice, maintaining a high
performance across different types of models.

## Experiment 2: Relationship with Sampling Characteristics

### Methods

We will be carrying out another simulation aimed at modulating the following
characteristics:

- **Model type**: linear or logistic.
- **"True" effect** (original regression coefficient from which data is drawn):
Can be 1 or 0 (no effect).
- **draws**: from 10 to 5000 by step of 5 (1000 iterations).
- **warmup**: Ratio of warmup iterations. from 1/10 to 9/10 by step of 0.1 (9
iterations).

We generated 3 datasets for each combination of these characteristics, resulting
in a total of `2 * 2 * 8 * 40 * 9 * 3 = 34560` Bayesian and frequentist models.
The code used for generation is avaible
[here](https://easystats.github.io/circus/articles/bayesian_indices.html)
(please note that it takes usually several days/weeks to complete).

```{r message=FALSE, warning=FALSE}
df <- read.csv("https://raw.github.com/easystats/circus/master/data/bayesSim_study2.csv")
```

### Results

#### Sensitivity to number of iterations

```{r, message=FALSE, warning=FALSE}

dat <- df
dat <- data_select(dat, select = c("iterations", "true_effect", "outcome_type", "beta", "Median", "Mean", "MAP"))
dat <- reshape_longer(
  dat,
  select = -c("iterations", "true_effect", "outcome_type"),
  names_to = "estimate",
  values_to = "value"
)
dat$temp <- as.factor(cut(dat$iterations, 5, labels = FALSE))

tmp <- lapply(split(dat, dat$temp), function(x) {
  x$iterations_group <- rep(round(mean(x$iterations), 1), times = nrow(x))
  return(x)
})

dat <- do.call(rbind, tmp)
dat <- data_filter(dat, value < 6)

ggplot(dat, aes(x = iterations_group, y = value, fill = estimate, group = interaction(estimate, iterations_group))) +
  geom_boxplot(outlier.shape = NA) +
  theme_classic() +
  scale_fill_manual(
    values = c("beta" = "#607D8B", "MAP" = "#795548", "Mean" = "#FF9800", "Median" = "#FFEB3B"),
    name = "Index"
  ) +
  ylab("Point-estimate of the true value 0\n") +
  xlab("\nNumber of Iterations") +
  facet_wrap(~ outcome_type * true_effect, scales = "free")
```

#### Sensitivity to warmup ratio

```{r, message=FALSE, warning=FALSE}
dat <- df
dat$warmup <- dat$warmup / dat$iterations
dat <- data_select(dat, select = c("warmup", "true_effect", "outcome_type", "beta", "Median", "Mean", "MAP"))
dat <- reshape_longer(
  dat,
  select = -c("warmup", "true_effect", "outcome_type"),
  names_to = "estimate",
  values_to = "value"
)
dat$temp <- as.factor(cut(dat$warmup, 3, labels = FALSE))

tmp <- lapply(split(dat, dat$temp), function(x) {
  x$warmup_group <- rep(round(mean(x$warmup), 1), times = nrow(x))
  return(x)
})

dat <- do.call(rbind, tmp)
dat <- data_filter(dat, value < 6)

ggplot(dat, aes(x = warmup_group, y = value, fill = estimate, group = interaction(estimate, warmup_group))) +
  geom_boxplot(outlier.shape = NA) +
  theme_classic() +
  scale_fill_manual(
    values = c("beta" = "#607D8B", "MAP" = "#795548", "Mean" = "#FF9800", "Median" = "#FFEB3B"),
    name = "Index"
  ) +
  ylab("Point-estimate of the true value 0\n") +
  xlab("\nNumber of Iterations") +
  facet_wrap(~ outcome_type * true_effect, scales = "free")
```


<!-- ## Experiment 3: Relationship with Priors Specification -->


## Discussion

Conclusions can be found in the [guidelines section](https://easystats.github.io/bayestestR/articles/guidelines.html) article.

# Suggestions

If you have any advice, opinion or such, we encourage you to let us know by
opening an [discussion thread](https://github.com/easystats/bayestestR/issues)
or making a pull request.
