---
title: "Example 1: Bayesian (General) Linear Models"
output: 
  github_document:
    toc: true
    fig_width: 10.08
    fig_height: 6
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, bayesian, posterior, test]
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{Example 1: Bayesian (General) Linear Models}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

This vignette can be referred to by citing the package:

- Makowski, D., Ben-Shachar M. S. \& Lüdecke, D. (2019). *Understand and Describe Bayesian Models and Posterior Distributions using bayestestR*. Available from https://github.com/easystats/bayestestR. DOI: [10.5281/zenodo.2556486](https://zenodo.org/record/2556486).

---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
options(digits=2)
```

Now that you've read the [**Get started**](https://easystats.github.io/bayestestR/articles/bayestestR.html) section, let's dive in the subtleties of Bayesian modelling using R.

## Loading the packages

Once you've [installed](https://easystats.github.io/bayestestR/articles/bayestestR.html) the necessary packages, we can load `rstanarm` (to fit the models), `bayestestR` (to compute useful indices) and `insight` (to access the parameters).

```{r message=FALSE, warning=FALSE}
library(rstanarm)
library(bayestestR)
library(insight)
```

## A Linear Model

### Fitting the model

We will start by fitting a simple model to test the relationship between **Petal.Length** (our predictor, - or independent, variable) and **Sepal.Length** (our response, - or dependent, variable) from the [`iris`](https://en.wikipedia.org/wiki/Iris_flower_data_set) dataset, included by default in R. 

Let's start by fitting the **frequentist** version of the model, just to have some **reference**:

```{r message=FALSE, warning=FALSE, eval=TRUE}
model <- lm(Sepal.Length ~ Petal.Length, data=iris)
summary(model)
```

In this model, the linear relationship between **Petal.Length** and **Sepal.Length** is **positive and significant** (beta = 0.41, t(148) = 21.65, p < .001).

Let's do the **Bayesian version**:

```{r message=FALSE, warning=FALSE, eval=FALSE}
model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, results='hide'}
library(rstanarm)
set.seed(333)

model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris)
```

### Extracting the posterior

You can see the sampling algorithm being run. Once it is done, let us extract the parameters (*i.e.*, the coefficients) of the model.

```{r message=FALSE, warning=FALSE}
posteriors <- insight::get_parameters(model)

summary(posteriors)  # Summary
```

As we can see, the parameters take the form of a long dataframe with two columns, corresponding to the **intercept** and **Petal.Length**. These columns contain the **posterior distributions** of these two parameters, *i.e.*, different plausible values.

#### About posterior *draws*

Let's look at the length of the posteriors.

```{r message=FALSE, warning=FALSE}
nrow(posteriors)  # Size (number of rows)
```

> **Why is the size 4000, and not more or less?**

First of all, these observations (the rows) are usually reffered to as **posterior draws**. The underlying idea is that the Bayesian sampling algorithm (such as **Monte Carlo Makov Chains - MCMC)**) will *draw* from the hidden true posterior distribution. Thus, it is through these posterior draws that we can estimate the underlying true posterior distribution. Thus, **the more draws you have, the better is your estimation of the posterior distribution**. But it also takes longer to compute.

If we look at the documentation (`?sampling`) for the rstanarm `"sampling"` algorithm used by default in the model above, we can see several parameters that influence the number of posterior draws. By default, there are **4** `chains` (you can see it as distinct sampling runs), all that will create **2000** `iter` (draws). However, only half of these iterations is kept, as half is used for `warmup` (the convergence of the algorithm). Which total is **`4 chains * (2000 iterations - 1000 warmup) = 4000`** posterior draws. We can change that, for instance:

```{r message=FALSE, warning=FALSE, eval=FALSE}
model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris, chains = 2, iter = 1000, warmup = 250)
 
nrow(insight::get_parameters(model))  # Size (number of rows)

```
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, results='hide'}
model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris, chains = 2, iter = 1000, warmup = 250)
nrow(insight::get_parameters(model))  # Size (number of rows)
```

In this case, we indeed have **`2 chains * (1000 iterations - 250 warmup) = 1500`** posterior draws. Let's keep our first model with the default setup.


#### Visualising the posterior distribution


Now that we've understood where do these values come from, let's look at them. We will start by visualizing (using the `ggplot2` package) the distribution of our parameter of interest, the effect of `Petal.Length`.


```{r message=FALSE, warning=FALSE}
library(ggplot2)

ggplot(posteriors, aes(x = Petal.Length)) +
  geom_density(fill = "orange")
```

This distribution represents the probability (the y axis) of different effects (the x axis). The central values are more probable than the extreme values. As you can see, this distribution ranges from about **0.35 to 0.50**, with the bulk of it being at around **0.41**.

> **Congrats! You've just described your posterior distribution.**

And this is at the heart of Bayesian analysis. We don't need *p*-values, *t*-values or degrees of freedom: **everything is there, under our nose**, within this **posterior distribution**.

As you can see, our description is consistent with the values of the frequentist regression (which had a beta of **0.41**). That is reassuring! We can now go ahead and precisely characterise this posterior distribution.

### Describing the Posterior

Unfortunately, it is often not practical to report the whole posterior distributions as graphs. We need to find a **consise way to summarise it**. We recommend to describe the posterior distribution with **3 elements**. A **point-estimate**, a one-value summary (such as the *beta* from the frequentist regression); a **credible interval**, representing the associated uncertainty and some **indices of significance**, giving information about the importance of this effect.


#### Point-estimate

#### Uncertainty

#### Effect significance



TBD.



