---
title: "Example 1: Linear Models"
output: 
  github_document:
    toc: true
    fig_width: 10.08
    fig_height: 6
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, bayesian, posterior, test]
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{Example 1: Linear Models}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

This vignette can be referred to by citing the package:

- Makowski, D., Ben-Shachar M. S. \& Lüdecke, D. (2019). *Understand and Describe Bayesian Models and Posterior Distributions using bayestestR*. Available from https://github.com/easystats/bayestestR. DOI: [10.5281/zenodo.2556486](https://zenodo.org/record/2556486).

---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
knitr::opts_chunk$set(comment=">")
options(digits=2)

set.seed(333)
```

Now that you've read the [**Get started**](https://easystats.github.io/bayestestR/articles/bayestestR.html) section, let's dive in the **subtleties of Bayesian modelling using R**.

## Loading the packages

Once you've [installed](https://easystats.github.io/bayestestR/articles/bayestestR.html) the necessary packages, we can load `rstanarm` (to fit the models), `bayestestR` (to compute useful indices) and `insight` (to access the parameters).

```{r message=FALSE, warning=FALSE}
library(rstanarm)
library(bayestestR)
library(insight)
```

## Simple linear model (*aka* a regression)

We will begin by a simple regression to test the relationship between **Petal.Length** (our predictor, - or *independent*, variable) and **Sepal.Length** (our response, - or *dependent*, variable) from the [`iris`](https://en.wikipedia.org/wiki/Iris_flower_data_set) dataset, included by default in R. 


### Fitting the model


Let's start by fitting the **frequentist** version of the model, just to have some **reference**:

```{r message=FALSE, warning=FALSE, eval=TRUE}
model <- lm(Sepal.Length ~ Petal.Length, data=iris)
summary(model)
```


In this model, the linear relationship between **Petal.Length** and **Sepal.Length** is **positive and significant** (beta = 0.41, t(148) = 21.6, p < .001). This means that for every increase of 1 in **Petal.Length** (the predictor), **Sepal.Length** (the response) increases by **0.41**. This can be visualised by plotting the observations:

```{r message=FALSE, warning=FALSE, eval=TRUE}
library(ggplot2)

ggplot(iris, aes(x=Petal.Length, y=Sepal.Length)) +
  geom_point() +
  geom_smooth(method="lm")
```


Let's fit a **Bayesian version** of the model:

```{r message=FALSE, warning=FALSE, eval=FALSE}
model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, results='hide'}
library(rstanarm)
set.seed(333)

model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris)
```

You can see the sampling algorithm being run. 

### Extracting the posterior

Once it is done, let us extract the parameters (*i.e.*, the coefficients) of the model.

```{r message=FALSE, warning=FALSE, eval=FALSE}
posteriors <- insight::get_parameters(model)

head(posteriors)  # Show the first 6 rows
```
```{r message=FALSE, warning=FALSE, echo=FALSE}
posteriors <- insight::get_parameters(model)

head(posteriors)  # Show the first 6 rows
```

As we can see, the parameters take the form of a long dataframe with two columns, corresponding to the **intercept** and the effect of **Petal.Length**. These columns contain the **posterior distributions** of these two parameters, *i.e.*, a set of different plausible values.

#### About posterior draws

Let's look at the length of the posteriors.

```{r message=FALSE, warning=FALSE}
nrow(posteriors)  # Size (number of rows)
```

> **Why is the size 4000, and not more or less?**

First of all, these observations (the rows) are usually reffered to as **posterior draws**. The underlying idea is that the Bayesian sampling algorithm (such as **Monte Carlo Makov Chains - MCMC**) will *draw* from the hidden true posterior distribution. Thus, it is through these posterior draws that we can estimate the underlying true posterior distribution. **The more draws you have, the better is your estimation of the posterior distribution**. But it also takes longer to compute.

If we look at the documentation (`?sampling`) for the rstanarm `"sampling"` algorithm used by default in the model above, we can see several parameters that influence the number of posterior draws. By default, there are **4** `chains` (you can see it as distinct sampling runs), that each create **2000** `iter` (draws). However, only half of these iterations is kept, as half is used for `warmup` (the convergence of the algorithm). The total is **`4 chains * (2000 iterations - 1000 warmup) = 4000`** posterior draws. We can change that, for instance:

```{r message=FALSE, warning=FALSE, eval=FALSE}
model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris, chains = 2, iter = 1000, warmup = 250)
 
nrow(insight::get_parameters(model))  # Size (number of rows)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
junk <- capture.output(model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris, chains = 2, iter = 1000, warmup = 250))
nrow(insight::get_parameters(model))  # Size (number of rows)
```

In this case, we indeed have **`2 chains * (1000 iterations - 250 warmup) = 1500`** posterior draws. Let's keep our first model with the default setup.


#### Visualising the posterior distribution


Now that we've understood where do these values come from, let's look at them. We will start by visualizing (using the `ggplot2` package) the distribution of our parameter of interest, the effect of `Petal.Length`.


```{r message=FALSE, warning=FALSE}
ggplot(posteriors, aes(x = Petal.Length)) +
  geom_density(fill = "orange")
```

This distribution represents the probability (the y axis) of different effects (the x axis). The central values are more probable than the extreme values. As you can see, this distribution ranges from about **0.35 to 0.50**, with the bulk of it being at around **0.41**.

> **Congrats! You've just described your posterior distribution.**

And this is at the heart of Bayesian analysis. We don't need *p*-values, *t*-values or degrees of freedom: **everything is there, under our nose**, within this **posterior distribution**.

As you can see, our description is consistent with the values of the frequentist regression (which had a beta of **0.41**). This is reassuring! We can now go ahead and **precisely characterise** this posterior distribution.

### Describing the Posterior

Unfortunately, it is often not practical to report the whole posterior distributions as graphs. We need to find a **consise way to summarise it**. We recommend to describe the posterior distribution with **3 elements**. A **point-estimate**, a one-value summary (such as the *beta* from the frequentist regression); a **credible interval**, representing the associated uncertainty and some **indices of significance**, giving information about the importance of this effect.


#### Point-estimate

**What single value can best represent my posterior distribution?**

Let's check the **mean**:

```{r message=FALSE, warning=FALSE}
mean(posteriors$Petal.Length)
```

This is close to the frequentist beta. But as we know, the mean is quite sensitive to outliers or extremes values. Maybe the **median** could be more robust?

```{r message=FALSE, warning=FALSE}
median(posteriors$Petal.Length)
```

Well, this is **close to the mean**. Maybe we could take the **mode**, the *peak* of the posterior distribution? In the Bayesian framework, this value is called the **Maximum A Posteriori (MAP)**. Let's see:

```{r message=FALSE, warning=FALSE}
map_estimate(posteriors$Petal.Length)
```

**They are all very close!** Let's visualise these values on the posterior distribution:

```{r message=FALSE, warning=FALSE}
ggplot(posteriors, aes(x = Petal.Length)) +
  geom_density(fill = "orange") +
  # The mean in blue
  geom_vline(xintercept=mean(posteriors$Petal.Length), color="blue", size=1) +
  # The median in red
  geom_vline(xintercept=median(posteriors$Petal.Length), color="red", size=1) +
  # The MAP in purple
  geom_vline(xintercept=map_estimate(posteriors$Petal.Length), color="purple", size=1)
```

Well, all these values give very similar results. Thus, **we will chose the median**, as this value has a direct meaning from a probabilistic perspective: **there is 50\% chance that the true effect is higher and 50\% chance that the effect is lower** (as it *cuts* the distribution in two equal parts).


#### Uncertainty

Now that the have a point-estimate, we have to **describe the uncertainty**. We could compute the range:

```{r message=FALSE, warning=FALSE}
range(posteriors$Petal.Length)
```

But does it make sense to include all these extreme values? Probably not. Thus, we will compute a [**credible interval**](https://easystats.github.io/bayestestR/articles/credible_interval.html). Long story short, it's kind of similar to a frequentist **confidence interval**, but it is easier to interpret and easier to compute. And it makes more sense.

We will compute this **credible interval** based on the [Highest Density Interval (HDI)](file:///C:/Users/Dom/Dropbox/RECHERCHE/N/easystats/bayestestR/docs/reference/hdi.html). It will give us the range containing the 90\% more probable effect values. **Note that we will use 90\% CIs instead of 95\% CIs as it is the case in the frequentist framework as the 90\% level gives more stable results**.

```{r message=FALSE, warning=FALSE}
hdi(posteriors$Petal.Length, ci=0.90)
```

Nice, so we can conclude that **the effect has 90\% chance of falling within the `[0.38, 0.44]` range**. We have just computed the two more important information to describe our effects. 

#### Effect significance

However, in many scientific fields, simply describing the effect is not enough. The readers also want to know if this effect ***means*** something. Is it important? Is it different from 0? In other words, **assessing the *significance* of the effect**. How can we do this?

Well, in this particular case, it is very eloquent: **all possible effect values (*i.e.*, the whole posterior distribution) are positive and superior to 0.35, which is already a lot**.

But still, we want some objective decision criterion, to say if **yes or no the effect is significant**.  We could do similarly to the frequentist framework, and see if the **Credible Interval** contains 0. Here, it is not the case, which means that our **effect is significant**. 

But this index is not very fine-grained, isn't it? **Can we do better? Yes.**


## A linear model with a categorical predictor

Let's take interest in the **weight of chickens**, depending on 2 **feed types**. We will start by selecting, from the `chickwts` dataset (also available in base R), the 2 feed types that are of interest for us (*because reasons*): **Meat meals** and **Sunflowers**.

### Data preparation and model fitting

```{r message=FALSE, warning=FALSE, eval=TRUE}
library(dplyr)

# We keep only rows for which feed is meatmeal or sunflower
data <- chickwts %>% 
  filter(feed %in% c("meatmeal", "sunflower"))
```


Let's run another Bayesian regression to predict the **Weight** with the **2 types of feed type**.

```{r message=FALSE, warning=FALSE, eval=FALSE}
model <- stan_glm(weight ~ feed, data=data)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, results='hide'}
model <- stan_glm(weight ~ feed, data=data)
```

### Posterior description


```{r message=FALSE, warning=FALSE, eval=TRUE}
posteriors <- insight::get_parameters(model)

ggplot(posteriors, aes(x=feedsunflower)) +
  geom_density(fill = "red")
```

This represents the **posterior distribution of the difference between meat meals and Sunflowers**. Seems that the difference is rather **positive**... Eating Sunflowers makes you more fat (if you're a chicken). **By how much?** Let us compute the **median** and the **CI**:

```{r message=FALSE, warning=FALSE, eval=TRUE}
median(posteriors$feedsunflower)
hdi(posteriors$feedsunflower)
```

It makes you fat by around `51` (the median). However, the uncertainty is quite high: **there is 90\% chance that the difference between the two feed types is between `6.21` and `88.67`.**

> **Is this effect different from 0?**

### ROPE Percentage

Testing wether this distribution is different from 0 doesn't make sense, as 0 is a single value (*and the probability that any distribution is different from a single value is infinite*). 

However, one way to assess **significance** could be to define an area around 0, which will consider as *equivalent* to zero (to no, - or negligible, effect). This is called the [**Region of Practical Equivalence (ROPE)**](https://easystats.github.io/bayestestR/articles/region_of_practical_equivalence.html), and is one way of testing the significance of parameters.

**How can we define this region?**


> ***Driing driiiing***

-- ***The easystats team speaking. How can we help?***

-- ***I am Prof Howard. An expert in chicks... I mean chickens. Just calling to let you know that based on my expert knowledge, an effect between -20 and 20 is negligible. Bye.***


Well, that's convenient. Now we know that we can define the ROPE as the `[-20, 20]` range. All effects within this range are considered as *null* (negligible). We can now compute the **proportion of the 90\% most probable values (the 90\% CI) which are not null**, *i.e.*, which are outside this range. 

```{r message=FALSE, warning=FALSE, eval=TRUE}
rope(posteriors$feedsunflower, range = c(-20, 20), ci=0.90)
```


**8.33\% of the 90\% CI can be considered as null**. Is that a lot? Based on our [**guidelines**](https://easystats.github.io/bayestestR/articles/guidelines.html), yes, it is too much. **Based on this particular definition of ROPE**, we conclude that this effect is not significant (the probability of being negligible is too high).

Altough, to be honest, I have **some doubts about this Prof Howard**. I don't really trust **his definition of ROPE**. Is there a more **objective** way of defining it?

**Yes.** One of the practice is for instance to use the **tenth (`1/10 = 0.1`) of the standard deviation (SD)** of the response variable, which can be considered as a "negligible" effect size.

```{r message=FALSE, warning=FALSE, eval=TRUE}
rope_value <- 0.1 * sd(data$weight)
rope_range <- c(-rope_value, rope_value)
rope_range
```

Let's redefine our ROPE as the region within the `[-6.2, 6.2]` range. **Note that this can be directly obtained by the `rope_range` function :)**

```{r message=FALSE, warning=FALSE, eval=TRUE}
rope_value <- rope_range(model)
rope_range
```

Let's recompute the **percentage in ROPE**:

```{r message=FALSE, warning=FALSE, eval=TRUE}
rope(posteriors$feedsunflower, range = rope_range, ci=0.90)
```

With this reasonable definition of ROPE, we observe that that the 90\% of the posterior distribution of the effect does **not** overlap with the ROPE. Thus, we can conclude that **the effect is significant**.


### Probability of Direction (pd)

Maybe we are not interested in whether the effect is not-negligible. Maybe **we just want to know if this effect is positive or negative**. In this case, we can simply compute the proportion of the posterior that is positive, no matter the "size" of the effect. 

```{r message=FALSE, warning=FALSE, eval=TRUE}
n_positive <- posteriors %>% 
  filter(feedsunflower > 0) %>% # select only positive values
  nrow() # Get length
n_positive / nrow(posteriors) * 100
```

We can conclude that **the effect is positive with a probability of 97.82\%**. We call this index the [**Probability of Direction (pd)**](https://easystats.github.io/bayestestR/articles/probability_of_direction.html). It can, in fact, be computed more easily with the following:

```{r message=FALSE, warning=FALSE, eval=TRUE}
p_direction(posteriors$feedsunflower)
```

Interestingly, it so happens that **this index is usually highly correlated with the frequentist *p*-value**. We could almost roughly infer the corresponding *p*-value with a simple transformation:

```{r message=FALSE, warning=FALSE, eval=TRUE}
pd <- 97.82
onesided_p <- 1 - pd / 100  
twosided_p <- onesided_p * 2
twosided_p
```

If we ran our model in the frequentist framework, we should approximately observe an effect with a *p*-value of `r round(twosided_p, digits=3)`. **Is that true?**




### Comparison to frequentist


```{r message=FALSE, warning=FALSE, eval=TRUE}
lm(weight ~ feed, data=data) %>% 
  summary()
```

The frequentist model tells us that the difference is **positive and significant** (beta = 52, p = 0.04). 

**Altough we arrived to a similar conclusion, the Bayesian framework allowed us to develop a more profound and intuitive understanding of our effect, and of the uncertainty of its estimation.**


## All with one function

And yet, I agree, it was a bit **tedious** to extract and compute all the indices. **But what if I told you that we can do all of this, and more, with only one function?**

> **Behold, `describe_posterior`!**

This function computes all of the afored mentioned indices, and can be run directly on the model:
```{r message=FALSE, warning=FALSE, eval=TRUE}
describe_posterior(model)
```

**Tada!** There we have it! The **median**, the **CI**, the **pd** and the **ROPE percentage**!

Understanding and describing posterior distributions is just one aspect of Bayesian modelling... **Are you ready for more?** [**Click here**](https://easystats.github.io/bayestestR/articles/example2_GLM.html) to see the next example.