---
title: "Example 1: Bayesian (General) Linear Models"
output: 
  github_document:
    toc: true
    fig_width: 10.08
    fig_height: 6
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, bayesian, posterior, test]
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{Example 1: Bayesian (General) Linear Models}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

This vignette can be referred to by citing the package:

- Makowski, D., Ben-Shachar M. S. \& Lüdecke, D. (2019). *Understand and Describe Bayesian Models and Posterior Distributions using bayestestR*. Available from https://github.com/easystats/bayestestR. DOI: [10.5281/zenodo.2556486](https://zenodo.org/record/2556486).

---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
options(digits=2)
```

Now that you've read the [**Get started**](https://easystats.github.io/bayestestR/articles/bayestestR.html) section, let's dive in the subtleties of Bayesian modelling using R.

## Loading the packages

Once you've [installed](https://easystats.github.io/bayestestR/articles/bayestestR.html) the necessary packages, we can load `rstanarm` (to fit the models) and `bayestestR` (to compute useful indices).

```{r message=FALSE, warning=FALSE}
library(rstanarm)
library(bayestestR)
library(insight)
```

## A Linear Model

### Fitting the model


```{r message=FALSE, warning=FALSE, eval=FALSE}
model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, results='hide'}
library(rstanarm)
set.seed(333)

model <- stan_glm(Sepal.Length ~ Petal.Length, data=iris)
```

### Extracting the posterior


```{r message=FALSE, warning=FALSE}
posteriors <- insight::get_parameters(model)
summary(posteriors)
```

As we can see, the parameters take the form of a long dataframe with two columns, corresponding to the **intercept** and **Petal.Length**. These columns contain the **posterior distributions** of these two parameters, *i.e.*, different plausible values. Let's look at the distribution of our parameter of interest, the effect of `Sepal.Length`.


```{r message=FALSE, warning=FALSE}
plot(density(posteriors$Petal.Length))
```


<!-- ### Correspondance with Frequentist Models -->

<!-- TO BE IMPROVED -->



